Make any stupid plot you want

### 2021-01-25 draft/sections/math.tex
fiber w/ 2D dimensions

\subsection{Data Space}
\label{sec:data}
We build on Butler's proposal of using fiber bundles as a common data representation format for visualization data\cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989} because fiber bundles are mathematical structures that are flexible enough express all the types of data described in section~\ref{sec:intro_data_structure}. The advantage of fiber bundles is they provide a way to encode the continuity in a dataset without making assumptions as to what that continuity is. In turn this representation of continuity can then be used to keep track of how the data fits together, for example if a visualization of a very large dataset calls for parallelization.

@@

\begin{figure}[ht!]
    \includegraphics[width=.5\linewidth]{figures/math/fiberbundle.png}

    \caption{ The fiber $F$ is a continuous plane encoding the set of all possible (time, temperature) values.  The base space $K$ expresses that the fibers lie in continous space and therefore the data is continuous. The mapping $\pi:E\rightarrow K$ from total space to base space is what defines this structure as a fiber bundle. The continuous function $\tau_1$ returns a record in each fiber for each point on $K$; the set of all values $\tau_1$ returns is a dataset that lies in this fiberbundle. \note{remove one of the sections from this figure/maybe kill this example}}
    \label{fig:data_base_space} 
\end{figure}
The fiber bundle in figure~\ref{fig:data_base_space} is a representation of continuous (temperature, time) data. The fiber bundle abstraction facilitates seperately representing the time and temperature variables in the fiber $F$ and how these variables are connected in the base space $K$.  The function $\tau_1$ returns a record (temperature, time) for ever point on the interval $K$ and there can be many $\tau$ functions in the fiber bundle.

\subsubsection{Variables: Fiber Space $F$}
\label{sec:data_fiber}
The fiber is a topological space that is the set of possible values of the data; the values themselves can be any dimension and type and have any continuity. We use Spivak's description of simplicial database schemas \cite{spivakSIMPLICIALDATABASES} as the basis of our fiber space because he binds the values in the fiber to variable names and types. Spivak constructs a set $U$ that is the disjoint union of all possible objects of the types $T\in DT$ that are in the datasets embedded in the fiber bundle. He then defines the single variable set $U_\sigma$ 
\begin{equation}
    \label{eq:data_types}
\begin{tikzcd}
    U_{\sigma} \arrow[r] \arrow[d, "\pi_{\sigma}"'] & U \arrow[d, "\pi"] \\
    C \arrow[r, "\sigma"']                          & \textbf{DT}       
\end{tikzcd}
\end{equation}
which is $U$ restricted to objects of type $T$ bound to variable name $c$. Given $U_{\sigma}$, the fiber for a one variable dataset is
\begin{equation}
    F = U_{\sigma(c)} = U_{T} 
\end{equation}
where $\sigma$ is the schema binding variable name $c$ to its datatype $T$. A dataset with multiple variables has a fiber that is the cartesian cross product of $U_{\sigma}$ applied to all the columns:
\begin{equation}
F = U_{T_1}\times \ldots U{T_i} \ldots\times U_{T_n}
\end{equation}

@@

which allows us to decouple $F$ into its component fibers $F_i$. For every variable
\begin{equation}
V_i = (c_i,\, T_i,\, U_{T_i})
\end{equation} 
we keep track of its name $c$, type $T$, and component in the fiber $U_{T}$.

\begin{figure}[ht!]
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/temp_2f.png}
        \label{fig:fiber_example_plane}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/temp_3f.png}
        \label{fig:fiber_example_cube}
    \end{subfigure}
    \label{fig:data_fiber_example}
    \caption{}.
\end{figure}

For example, the data in figure~\ref{fig:fiber_example_plane} is a set of temperature measurements in \textdegree C and the times those measurements were taken at. Temperature values are floating point numbers that come from the set of real numbers $U_{float} = \mathbb{R}$ and times that are always positive $U_{datetime}= \mathbb{R}^+$, so the fiber is 
\begin{equation}
    F = \mathbb{R} \times \mathbb{R}^{+}
\end{equation} 
where $V_0 = (temperature,\, \texttt{float},\, \mathbb{R})$ and $V_1 = (time,\, \texttt{datetime},\, \mathbb{R}^+)$
In figure~\ref{fig:fiber_example_cube}, a third variable wind is added to the dataset. This wind variable is of type \texttt{wind} and has two components speed and direction $\{(m,d) \in \mathbb{R}^{2} \mid  0\leq m,\, 0 \leq d \leq 360\}$ . The new variable is added to the fiber
\begin{equation}
    F = \mathbb{R} \times \mathbb{R} \times \mathbb{R}^2
\end{equation} 
such that $V_2 = (wind,\, \texttt{wind},\, \mathbb{R}^{2})$

\subsubsection{Measurement Scales: Monoid Actions}
\label{sec:data_monoid}
We formally describe the set of allowable operations in the fiber because a transformation on the data side needs to be reflected in the visual representation, discussed in more detail in section~\ref{sec:artist_nu}. We generalize this set of allowable transformations to the monoid actions $M_i$ on $F_i$. A monoid \cite{Monoid2021} $M_i$ is a set with an associative binary operator $\ast:M_i \times M_i\rightarrow M_i$. A monoid has an identity element $e\in M_i$ such that $e\ast a= a \ast e = a$ for all $a \in M_i$. A left monoid action \cite{SemigroupAction2021,ActionNLab} of $M_i$ is a set $U_{T_i}$ with an action $\bullet: M\times U_{T_i} \rightarrow U_{T_i}$ with the properties:
\begin{align*}
    \textbf{associativity}\;& \text{for all } f,g \in M_i \text{ and } x\in U_{T_i}, f\bullet(g\bullet x) = (f\ast g) \bullet x\\
    \textbf{identity}\;& \text{for all } x\in U_{T_i}, e\in M_i,  e\bullet x = x 
\end{align*}
As with the fiber $F$, the total monoid space $M$ is the cartesian product
\begin{equation}
M = M_{0} \times \ldots \times M_{i}\times \ldots \times\ldots M_{n}
\end{equation}
of the monoid $M_{i}$ on each $F_{i}\in F$. We expand our definition of variables 
\begin{equation}
        V_i = (c_i,\, T_i,\, U_{T_i}\, M_i)
\end{equation}
to include the monoid actions that act on $U_{T_i}$.

Steven's defined the measurement scales \cite{stevensTheoryScalesMeasurement1946,leaFormalizationMeasurementScale} in terms of the monoid actions permissible on the measurments; nominal data is permutable, ordinal data is monotonic, interval data is translatable, and ratio data is scalable \cite{weissteinSimilarityTransformation}.

@@

\subsubsection{continuity: Base Space $K$} 
%%fold butlder
\label{sec:data_base}
The base space $K$ provides a way to encode how the records are connected to each other; for example the GHCN data can be expressed as discrete observations, many timeseries, maps, a network of stations, etc.

\begin{figure}[ht!]
    \includegraphics[width=.5\textwidth]{figures/math/k_different_types.png}
    \label{fig:base_space_types}
    \caption{The topological base space $K$ encodes the connectivity of the data space, for example if the data is independent points or a map or on a sphere}
\end{figure}

$K$ has no knowledge of the records. As illustrated in figure~\ref{fig:base_space_div}, $K$ is more of an indexing space into $E$ that describes the structure of $E$ and itself can be any number of dimensions.

\begin{figure}[ht!]
    \label{fig:base_space_div}
    \includegraphics[width=.5\linewidth]{figures/math/k_qspace.png}
    \caption{ The base space $E$ is divided into fiber segments $F$. The base space $K$ acts as an index into the fibers.}
\end{figure}

Formally $K$ is the quotient space \cite{QuotientSpaceTopology2020} of $E$, meaning it is the finest space such that every $k \in K$ has a corresponding fiber $F_k$ in $E$. In figure~\ref{fig:base_space_div}, $E$ is a rectangle divided by vertical fibers $F$, so the minimal $K$ for which there is always a mapping $\pi: F\rightarrow K$ is the line.

@@

\subsubsection{Values: Sections $\tau$}
\label{sec:data_section}
\begin{figure}[ht!]
    \label{fig:base_space_div}
    \includegraphics[width=.5\linewidth]{figures/math/k_qspace.png}
    \caption{ A section $\tau$ is a mapping from $K$ to $F$ that returns a record for each $k \in K$.}
\end{figure}

@@

\subsubsection{{Example}}
\begin{figure}[ht!]
    \includegraphics[width=.4\linewidth]{figures/math/render.png}
    \caption{}
    \label{fig:render}
\end{figure}

@@

\subsubsection{Visual Channels}
\label{sec:artist_nu}
$\nu: E \rightarrow V$ is an equivariant map such that there is a homomorphism from left monoid actions on $E_{i}$ to left monoid actions on $V_{i}$ where $i$ identifies a field in the fiber. $E_i$ and $V_{i}$ each contain a set of values as defined in $F$ and $P$ respectively. A validly constructed $\nu$ is one where the  diagram 
\begin{equation}
    \label{eq:nu_categorical}
\begin{tikzcd}
    E_i \arrow[r] \arrow[r, "\nu_i"] \arrow[d, "m_e"'] & V_i \arrow[d, "m_v"] \\
    E_i \arrow[r, "\nu_i"]                           & V_i               
\end{tikzcd}
\end{equation}
commutes such that $\nu_i(m_e(E_i)) = m_v(\nu_i(E_i))$.

\paragraph{Example: Ordering}
To preserve ordering of elements in $E_i$, $\nu$ must be a monotonic function such that given $e_1, e_2 \in E_{i}$
\begin{equation}
\text{ if } e_1 \leq e_2 \text{ then } \nu(e_1) \leq \nu(e_2)
\end{equation}

\paragraph{Example: Translation}
According to Stevens, interval data is a set with general linear group actions \cite{stevensTheoryScalesMeasurement1946,leaFormalizationMeasurementScale}. Position is a visual variable that can support translation 
\begin{equation}
\nu(x + c) = \nu(x) + \nu(c)
\end{equation}

@@

\subsubsection{Assembling Marks}
\label{sec:artist_q}
\begin{figure}[ht!]
    \includegraphics[width=\textwidth]{figures/math/path_of_q}
    \label{fig:artist_q}
    \caption{The $\nu$ functions convert data $E$ to visual $V$. $Q$ assembles the different types of visual parameters $V_{i}$ into a graphic in $H$. $Q\circ\mu(\xi^{-1}J)$ forms a visual mark by applying $Q$ to a region mapped to connected components $J \subset K$.}
    %%maybe change to \tau, \mu, \rho  
\end{figure}

@@

\begin{figure}[ht!]
    \includegraphics[width=\textwidth]{figures/math/diff_type_q.png}
    \label{fig:artist_mark_change}
\end{figure}

@@

\paragraph{Q: scatter plot}\mbox{} \\
\label{sec:artist_example_scatter}
\begin{figure}[ht!]
    \includegraphics[width=\textwidth]{figures/math/scatter.png}
    \label{fig:artist_scatter}
\end{figure}

@@

\paragraph{Q: line plot}\mbox{} \\
\label{sec:artist_example_line}
\begin{figure}[ht!]
    \includegraphics[width=\textwidth]{figures/math/line.png}
    \label{fig:artist_line}
\end{figure}

@@

\paragraph{Q: heatmap}\mbox{} \\
\label{sec:artist_example_heatmap}
\begin{figure}[ht!]
    \includegraphics[width=\textwidth]{figures/math/heatmap.png}
    \label{fig:artist_heatmap}
\end{figure}
The heatmap in figure~\ref{fig:artist_heatmap} 
\begin{equation}
Q(xpos, ypos, color)
\end{equation}
has in the simple case a direct lookup into $K$ to obtain the $\mu = (x,y,c)$ values that are mapped into 
(arrays are upside down, x is y, is x) - $\xi$ is about translating indices from data to visual alignment
\begin{align}
D_{RGB} &= color(\xi(\alpha_1, \alpha_2))
D_x & = xpos(\xi(\alpha_1, \alpha_2))
D_y &= ypos(\xi(\alpha_1, \alpha_2))
\end{align}
through $\rho$.

@@

\begin{figure}[ht!]
    \includegraphics{figures/math/simplex.png}
    \caption{Simplices can encode the connectivity of the data, from fully disconnected (0 simplex) records to all records are connected to at least 3 others}
    \label{fig:triangle_simplex}
\end{figure}

@@

\subparagraph{Example: Mobius Strip}\mbox{}\\
\begin{figure}[ht!]
    \label{fig:data_base_transition}
    \includegraphics[width=\textwidth]{figures/math/transition_functions.png}
    \caption{Many non-trivial spaces can be made locally trivial by dividing $E$ into locally trivial subspaces and defining transition functions between the edges on K for how to glue the two subspaces such that the sigmas are continuous.}
\end{figure}

@@

\subparagraph{Example: Torus}\mbox{}\\
\begin{figure}[ht!]
    \label{fig:triangle_torus}
    \includegraphics[width=\textwidth]{figures/math/triangle_torus.png}
    \caption{Representation of the base space $K$ of a toridal total space $E$ as two connected 2-simplices.}
\end{figure}

>>>>

\subsection{Data Space}
\label{sec:data}
We build on Butler's proposal of using fiber bundles as a common data representation format for visualization data\cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989} because fiber bundles are mathematical structures that are flexible enough express all the types of data described in section~\ref{sec:intro_data_structure}.

@@

\subsubsection{Variables: Fiber Space $F$}
\label{sec:data_fiber}
The fiber is a topological space that is the set of possible values of the data; the values themselves can be any dimension and type and have any continuity. We use Spivak's description of simplicial database schemas \cite{spivakSIMPLICIALDATABASES} as the basis of our fiber space because he binds the components of the fiber to variable names and types. Spivak constructs a set $U$ that is the disjoint union of all possible objects of types $\{T_0, \ldots, T_n\} \in DT$ where $DT$ are the data types of the variables in the dataset. He then defines the single variable set $U_\sigma$ 
\begin{equation}
    \label{eq:data_types}
\begin{tikzcd}
    U_{\sigma} \arrow[r] \arrow[d, "\pi_{\sigma}"'] & U \arrow[d, "\pi"] \\
    C \arrow[r, "\sigma"']                          & \textbf{DT}       
\end{tikzcd}
\end{equation}
which is $U$ restricted to objects of type $T$ bound to variable name $c$. Given $U_{\sigma}$, the fiber for a one variable dataset is
\begin{equation}
    F = U_{\sigma(c)} = U_{T} 
\end{equation}
where $\sigma$ is the schema binding variable name $c$ to its datatype $T$. A dataset with multiple variables has a fiber that is the cartesian cross product of $U_{\sigma}$ applied to all the columns:
\begin{equation}
F = U_{T_1}\times \ldots U{T_i} \ldots\times U_{T_n}
\end{equation}

@@

which allows us to decouple $F$ into its component fibers $F_i$. For every variable
\begin{equation}
V_i = (c_i,\, T_i,\, U_{T_i})
\end{equation} 
we keep track of its name $c$, type $T$, and values $U_{T}$.

\begin{figure}[H]
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/temp_2f.png}
        \caption{}
        \label{fig:fiber_example_plane}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/temp_3f.png}
        \caption{}
        \label{fig:fiber_example_cube}
    \end{subfigure}
    \caption{These two datasets have the same base space $K$, but figure~\ref{fig:fiber_example_plane} has fiber  $F=\mathbb{R}\times\mathbb{R}^+$ which is (time, temperature) while figure~\ref{fig:fiber_example_cube} has fiber $\mathbb{R}^{+}\times\mathbb{R}^2$ which is (time, wind=(speed, direction))}
    \label{fig:data_fiber_example}
\end{figure}

For example, the data in figure~\ref{fig:fiber_example_plane} is a pair of times and \textdegree C temperature measurements taken at those times. Time is a positive number ot type \textt{detetime} which can be resolved to positive floats $U_{\texttt{datetime}}= \mathbb{R}^+$. Temperature values are real numbers $U_{\texttt{float}} = \mathbb{R}$. The fiber is 
\begin{equation}
    F =  \mathbb{R}^{+} \times \mathbb{R} 
\end{equation} 
where $V_0 = (time,\, \texttt{datetime},\, \mathbb{R}^+)$ and $V_1 = (temperature,\, \texttt{float},\, \mathbb{R})$. In figure~\ref{fig:fiber_example_cube}, temperature is replaced with wind. This wind variable is of type \texttt{wind} and has two components speed and direction $\{(s,d) \in \mathbb{R}^{2} \mid  0\leq s,\, 0 \leq d \leq 360\}$ . Therefore, the fiber is 
\begin{equation}
    F = \mathbb{R}^{+} \times \mathbb{R}^2
\end{equation} 
such that $V_1 = (wind,\, \texttt{wind},\, \mathbb{R}^{2})$

\subsubsection{Measurement Scales: Monoid Actions}
\label{sec:data_monoid}
We formally describe the set of allowable operations in the fiber because a transformation on the data side needs to be reflected in the visual representation, discussed in more detail in section~\ref{sec:artist_nu}. Steven's described the measurement scales\cite{stevensTheoryScalesMeasurement1946,leaFormalizationMeasurementScale} in terms of these transformations: nominal data is permutable, ordinal data is monotonic, interval data is translatable, and ratio data is scalable \cite{weissteinSimilarityTransformation}.

We generalize the set of allowable transformations to the monoid actions $M_i$ on $F_i$. A monoid \cite{Monoid2021} $M_i$ is a set with an associative binary operator $\ast:M_i \times M_i\rightarrow M_i$. A monoid has an identity element $e\in M_i$ such that $e\ast a= a \ast e = a$ for all $a \in M_i$. A left monoid action \cite{SemigroupAction2021,ActionNLab} of $M_i$ is a set $U_{T_i}$ with an action $\bullet: M\times U_{T_i} \rightarrow U_{T_i}$ with the properties:
\begin{align*}
    \textbf{associativity}\;& \text{for all } f,g \in M_i \text{ and } x\in U_{T_i}, f\bullet(g\bullet x) = (f\ast g) \bullet x\\
    \textbf{identity}\;& \text{for all } x\in U_{T_i}, e\in M_i,  e\bullet x = x 
\end{align*}
As with the fiber $F$, the total monoid space $M$ is the cartesian product
\begin{equation}
M = M_{0} \times \ldots \times M_{i}\times \ldots \times\ldots M_{n}
\end{equation}
of the monoid $M_{i}$ on each $F_{i}\in F$. We expand our definition of variables 
\begin{equation}
        V_i = (c_i,\, T_i,\, U_{T_i}\, M_i)
\end{equation}
to include the monoid actions that act on $U_{T_i}$.

@@

\subsubsection{Continuity: Base Space $K$} 
\label{sec:data_base}
The advantage of fiber bundles is they provide a way to encode the continuity in a dataset as the basespace $K$ without making assumptions as to what that continuity is. In turn this representation of continuity can then be used to keep track of how the data fits together, for example if a visualization of a very large dataset calls for parallelization.

\begin{figure}[H]
    \includegraphics[width=.5\textwidth]{figures/math/k_different_types.png}
    \caption{The topological base space $K$ encodes the connectivity of the data space, for example if the data is independent points or a map or on a sphere}
    \label{fig:base_space_types}
\end{figure}
As illustrated in figure~\ref{fig:base_space_types}, $K$ is akin to an indexing space into $E$ that describes the structure of $E$.  $K$ can have any number of dimensions and be continuous or discrete.

\begin{figure}[H]
    \includegraphics[width=.5\linewidth]{figures/math/k_qspace.png}
    \caption{The base space $E$ is divided into fiber segments $F$. The base space $K$ acts as an index into the records in the fibers.
    \note{this figure might be good all the way up top to lay out the components of fb}}
    \label{fig:base_space_div}
\end{figure}

Formally $K$ is the quotient space \cite{QuotientSpaceTopology2020} of $E$, meaning it is the finest space\cite{aurouxMath131Introduction} such that every $k \in K$ has a corresponding fiber $F_k$ in $E$\cite{QuotientSpaceTopology2020}. In figure~\ref{fig:base_space_div}, $E$ is a rectangle divided by vertical fibers $F$, so the minimal $K$ for which there is always a mapping $\pi: F\rightarrow K$ is the line.

@@

\begin{figure}[ht!]
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/temp_1k.png}
        %% add box around neighboring P and Map
        \caption{}
        \label{fig:base_example_discrete}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/temp_2k.png}
        \caption{}
        \label{fig:base_example_continuous}
    \end{subfigure}
    \caption{These two datasets have the same (time, temperature) fiber, but in figure~\ref{fig:base_example_line} the records are discrete, while in figure~\ref{fig:base_example_plane} the records are sampled from the continuous interval $K$
    \note{revamp figure: F=Plane, k1 = dots, k2=line}}
    \label{fig:base_example}
\end{figure}

The datasets in figure~\ref{fig:base_example} have the same fiber of (temperature, time) which is $F=\mathbb{R}\times \mathbb{R}^+$. In figure~\ref{fig:base_example_discrete} the fibers lie over discrete $K$ such that the records in the datasets in the fiber bundles are discrete. The same fiber in figure~\ref{fig:base_example_continuous} lies over a continuous interval $K$ such that the records are samples from a continuous function defined on $K$.

\subsubsection{Data: Sections $\tau$}
\label{sec:data_section}
\begin{figure}[H]
    \label{fig:base_space_div}
    \includegraphics[width=.5\linewidth]{figures/math/k_qspace.png}
    \caption{ A section $\tau$ is a mapping from $K$ to $F$ that returns a record for each $k \in K$.}
\end{figure}

@@

\subsubsection{{Example}}
\begin{figure}[H]
    \includegraphics[width=.4\linewidth]{figures/math/render.png}
    \caption{}
    \label{fig:render}
\end{figure}

@@

\subsubsection{Visual Channels}
\label{sec:artist_nu}
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/equivariance_nu.png}
    \caption{An equivariant $\nu$ preserves monoid actions. Here $\nu$ maps the words to the emojis and is constructed such that a permutation of the words is reflected in a permutation of the symbols. \note{will switch to a figure table thing but this is the idea}}
    \label{fig:artist_nu}
\end{figure}
The expectation of a visual transformation is that it will convert data to visual variables in a consistent manner. For example in figure~\ref{fig:artist_nu}, the $\nu$ maps words to emojies. Our constraint on $\nu$ is that if we shuffle the words, the emojis should be shuffled such that the map defined by $\nu$ is still valid. Formally, this means that $\nu: E \rightarrow V$ is an equivariant map such that there is a homomorphism from left monoid actions on $E_{i}$ to left monoid actions on $V_{i}$ where $i$ identifies a field in the fiber. $E_i$ and $V_{i}$ each contain a set of values as defined in $F$ and $P$ respectively. A validly constructed $\nu$ is one where the  diagram 
\begin{equation}
    \label{eq:nu_categorical}
\begin{tikzcd}
    E_i \arrow[r] \arrow[r, "\nu_i"] \arrow[d, "m_e"'] & V_i \arrow[d, "m_v"] \\
    E_i \arrow[r, "\nu_i"]                           & V_i               
\end{tikzcd}
\end{equation}
commutes such that $\nu_i(m_e(E_i)) = m_v(\nu_i(E_i))$.

\paragraph{Example: Ordering}
To preserve ordering of elements, $\nu$ must be a monotonic function such that given $e_1, e_2 \in E_{i}$
\begin{equation}
\text{ if } e_1 \leq e_2 \text{ then } \nu(e_1) \leq \nu(e_2)
\end{equation}

\paragraph{Example: Translation}
A $\nu$ is equivariant under translation monoid actions if 
\begin{equation}
\nu(x + c) = \nu(x) + \nu(c)
\end{equation}

@@

\subsubsection{Assembling Marks}
\label{sec:artist_q}
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/path_of_q}
    \label{fig:artist_q}
    \caption{The $\nu$ functions convert data $E$ to visual $V$. $Q$ assembles the different types of visual parameters $V_{i}$ into a graphic in $H$. $Q\circ\mu(\xi^{-1}J)$ forms a visual mark by applying $Q$ to a region mapped to connected components $J \subset K$.}
    %%maybe change to \tau, \mu, \rho  
\end{figure}

@@

\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/diff_type_q.png}
    \label{fig:artist_mark_change}
\end{figure}

@@

\paragraph{Q: scatter plot}\mbox{} \\
\label{sec:artist_example_scatter}
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/scatter.png}
    \label{fig:artist_scatter}
\end{figure}

@@

\paragraph{Q: line plot}\mbox{} \\
\label{sec:artist_example_line}
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/line.png}
    \label{fig:artist_line}
\end{figure}

@@

\paragraph{Q: heatmap}\mbox{} \\
\label{sec:artist_example_heatmap}
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/heatmap.png}
    \label{fig:artist_heatmap}
\end{figure}
The heatmap in figure~\ref{fig:artist_heatmap} 
\begin{equation}
Q(xpos, ypos, color)
\end{equation}
has in the simple case a direct lookup into $K$ to obtain the $\mu = (x,y,c)$ values that are mapped into 
(arrays are upside down, x is y, is x) - $\xi$ is about translating indices from data to visual alignment
\begin{align}
D_{RGB} &= color(\xi(\alpha_1, \alpha_2))
D_x & = xpos(\xi(\alpha_1, \alpha_2))
D_y &= ypos(\xi(\alpha_1, \alpha_2))
\end{align}
through $\rho$.

@@

\begin{figure}[H]
    \includegraphics{figures/math/simplex.png}
    \caption{Simplices can encode the connectivity of the data, from fully disconnected (0 simplex) records to all records are connected to at least 3 others}
    \label{fig:triangle_simplex}
\end{figure}

@@

\subparagraph{Example: Mobius Strip}\mbox{}\\
\begin{figure}[H]
    \label{fig:data_base_transition}
    \includegraphics[width=\textwidth]{figures/math/transition_functions.png}
    \caption{Many non-trivial spaces can be made locally trivial by dividing $E$ into locally trivial subspaces and defining transition functions between the edges on K for how to glue the two subspaces such that the sigmas are continuous.}
\end{figure}

@@

\subparagraph{Example: Torus}\mbox{}\\
\begin{figure}[H]
    \label{fig:triangle_torus}
    \includegraphics[width=\textwidth]{figures/math/triangle_torus.png}
    \caption{Representation of the base space $K$ of a toridal total space $E$ as two connected 2-simplices.}
\end{figure}

### 2021-01-25 draft/sections/scraps.tex
fiber w/ 2D dimensions

>>>>

\begin{figure}[ht!]
    \includegraphics[width=\textwidth]{figures/math/rotation_actions.png}
    \label{fig:data_monoid_rotation}
    \caption{The set of rotation actions [turn 0, 90, 180] on the star produce a star (closed), include a rotation that does not change the orientation (identity), and can be done in any order to produce the same final orientation (closure); therefore the rotations are a set of monoid actions.
    %% dataset that is symmetric, fiber is x,y, rotatation action
    %% rotate + plot or plot + rotate
    %%emoji shuffle 
    }  
\end{figure}

\begin{figure}[ht!]
    \includegraphics[width=.5\linewidth]{figures/math/fiberbundle.png}

    \caption{ The fiber $F$ is a continuous plane encoding the set of all possible (time, temperature) values.  The base space $K$ expresses that the fibers lie in continous space and therefore the data is continuous. The mapping $\pi:E\rightarrow K$ from total space to base space is what defines this structure as a fiber bundle. The continuous function $\tau_1$ returns a record in each fiber for each point on $K$; the set of all values $\tau_1$ returns is a dataset that lies in this fiberbundle. \note{remove one of the sections from this figure/maybe kill this example}}
    \label{fig:data_base_space} 
\end{figure}
The fiber bundle in figure~\ref{fig:data_base_space} is a representation of continuous (temperature, time) data. The fiber bundle abstraction facilitates seperately representing the time and temperature variables in the fiber $F$ and how these variables are connected in the base space $K$.  The function $\tau_1$ returns a record (temperature, time) for ever point on the interval $K$ and there can be many $\tau$ functions in the fiber bundle.
